|

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
cleaned up common factors section

```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted plot."}

@@

XXX DC will clean this up

Four distributions of fitted values are also considered by using $X_{raw}$ follows different distributions: (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$.

Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.

>>>>

```{r residual-plot-common-departures, fig.width = 10, fig.height = 2.5, fig.cap = "Example residual vs fitted value plots: (A) classically good looking residuals, (B) non-linear pattern indicates that the model has not captured a non-linear association, (C) heteroskedasticity indicating that variance around the fitted model is not uniform, and (D) non-normality where the residual distribution is not symmetric around 0. The latter pattern might best be assessed using a univariate plot of the residuals, but patterns B and C need to be assessed using a residual vs fitted value plot."}

@@

Fitted values are a function of the independent variables, and the distribution of the observed values affects the distribution of the fitted values. In the best case scenario the fitted values will have a uniform distribution, which means that there is even coverage of possible observed values across all of the predictors. This is not always present in the collected data. Sometimes the fitted values are discrete because one or more predictors were measured discretely. The distribution may be relatively Gaussian, reflecting a linear combinatio of many predictors, adhering to the Central Limit Theorem. It is also common to see a skewed distribution of fitted values, if one or more of the predictors has a skewed distribution. This latter problem is usually corrected before modeling using a variable transformation. Our simulation asesses this by using four different distributions to represent the fitted values ($X_{raw}$): (1) $U(-1, 1)$, (2) $N(0, 0.3^2)$, (3) $lognormal(0, 0.6^2)/3$ and (4) $u\{1, 5\}$. We would expect that the best reading of residual plots occurs when the fitted values are uniformly distributed.

<!--Different distributions of fitted values help enriching the pool of visual patterns as illustrated in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot.-->

Three different sample sizes are used, $n=50, 100, 300$ across the experiments. We would expect considerable variation in the signal strength in the simulated data plots with smaller $n$. A sample size of 300 is typically enough for structure to be visible in a scatterplot reliably.
